---
title: "p8131_hw6_mp3745"
author: "Matthew Parker"
date: '`r format(Sys.time(), "%Y-%m-%d")`'
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(nnet)
library(MASS)
library(pscl)
library(nlme)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)
options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)
scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
theme_set(theme_minimal() + theme(legend.position = "bottom"))
```


## Problem 1

Handwritten


## Problem 2

Read in the data
```{r}
dental = read.table("./data/HW6-dental.txt", header = TRUE) %>% 
  janitor::clean_names() %>% 
  as_tibble() %>% 
  mutate(
    gender = recode(gender, '0' = "F", '1' = "M"),
    gender = as_factor(gender)
  ) 
```

#### (1)

Make a spaghetti plot to infer how distance is aô°€ected by age and gender
```{r}
dental %>% 
  ggplot(aes(x = age, y = distance, color = gender, group = child)) +
  geom_line()
```

Based on the above plot, distance appears to increase with age and distance tends to be higher for males compared to females.


#### (2)

Handwritten


#### (3)
###### a.

Fit a model with compound symmetry covariance
```{r}
# Fit
compound = gls(distance ~ gender + age,
               data = dental,
               correlation = corCompSymm(form = ~ 1 | child),
               weights = varIdent(form = ~ 1| age),
               method = "REML")

# Summary
summary(compound)

# Sigma for compound
compound_sigma = sigma(compound)

(1 + compound$modelStruct$varStruct[1]) * compound_sigma
```


###### b.

Fit a model with exponential covariance
```{r}
# Fit
exponential = gls(distance ~ gender + age,
               data = dental,
               correlation = corExp(form = ~ 1 | child),
               weights = varIdent(form = ~ 1| age),
               method = "REML")

# Summary
summary(exponential)

# Sigma for exponential
exponential_sigma = sigma(exponential)
```


###### c.

Fit a model with autoregressive covariance
```{r}
# Fit
autoregressive = gls(distance ~ gender + age,
               data = dental,
               correlation = corAR1(form = ~ 1 | child),
               weights = varIdent(form = ~ 1| age),
               method = "REML")

# Summary
summary(autoregressive)

# Sigma for autoregressive
autoregressive_sigma = sigma(autoregressive)
```


Put together dataframe to summarize coefficient parameter estimates and covariance estimates from fits with different covariance patterns
```{r}
fit_summary = tibble(
  fit = c("compound", "exponential", "autoregressive"),
  intercept = c(compound$coefficients[1], exponential$coefficients[1], autoregressive$coefficients[1]),
  gender = c(compound$coefficients[2], exponential$coefficients[2], autoregressive$coefficients[2]),
  age = c(compound$coefficients[3], exponential$coefficients[3], autoregressive$coefficients[3]),
  cov_age_8 = c(compound_sigma, exponential_sigma, autoregressive_sigma),
  cov_age_10 = c(compound_sigma * (1 + compound$modelStruct$varStruct[1]), 
                 exponential_sigma * (1 + exponential$modelStruct$varStruct[1]), 
                 autoregressive_sigma * (1 + autoregressive$modelStruct$varStruct[1])),
  cov_age_12 = c(compound_sigma * (1 + compound$modelStruct$varStruct[2]), 
                 exponential_sigma * (1 + exponential$modelStruct$varStruct[2]), 
                 autoregressive_sigma * (1 + autoregressive$modelStruct$varStruct[2])),
  cov_age_14 = c(compound_sigma * (1 + compound$modelStruct$varStruct[3]), 
                 exponential_sigma * (1 + exponential$modelStruct$varStruct[3]), 
                 autoregressive_sigma * (1 + autoregressive$modelStruct$varStruct[3]))
)

fit_summary %>% 
  knitr::kable()
```

The coefficient parameter and covariance estimates are nearly identical for exponential and autoregressive. The coeffecient estimate for gender is smaller for compound, but age estimate is larger. The covariance estimates at age 8 and 10 are smaller for the compound, but the covariance estimate at age 14 is higher.